Zohmg---a large scale data warehouse for aggregated time-series-based data. [DRAFT]

### 2009-05-05 doc/report/msc-report.tex
doc/report/msc-report.tex: + added section on HDFS. + %s/OHM/Zohmg/g + minor changes here and there.

Each column-family is stored as a single map-file in hbase.

@@

@@

Hadoop is an open source software implementation of Google's MapReduce-framework described in \cite{mapreduce}. Hadoop is written in Java and is part of the Apache Software Foundation.  It is a fairly mature piece of software and is used in production at numerous companies.

@@

MapReduce works by splitting the input, which commonly is in the size range of many gigabytes, into n parts and lets each node of the computing cluster work on one or more parts. The mapper on each node emits key-value pairs which are fed into the reduce phase. The final output is collated, usually on a network file system.

@@

@@

@@

HBase is an open source implementation of Google's BigTable data store described in \cite{bigtable}. It is a sub-project of Hadoop, written in Java and a part of the Apache Software Foundation. HBase is still in its infancy and few organizations use it for mission-critical applications.

@@

The typical use case for Zohmg is to read one or more qualifiers, which correspond to one or more points in n-space, over a range of time.  Therefor, the keys are made up of the time and unit of the measurement, and the qualifier queried for represents the point in n-space. More on this in the data model section.

@@

\subsubsection{Python}

@@

Python Setuptools is a framework---built on top of the Python module
distutils---for basic package management of Python software. Amongst other
things it creates and installs Python Eggs\footnote{A Zip file which
bundles Python software.}.

@@

As stated in chapter 3, Zohmg is built upon existing tools and software. It
uses Apache Hadoop and HBase, free implementations of Google's MapReduce
and BigTable respectively, both implemented in Java.

@@

Data is stored in Apache HBase.

Exporting data from Zohmg is done via a RESTful interface which dumps JSON.
It is possible to transform the stored data with user defined transformers.

OHM is implemented in Python.

@@

Certain paths need to be known by OHM in order to function, these are set
in the config/environment.py file.

@@

The user defines a mapper which OHM wraps, interpreting the output from the
mapper program and storing it in HBase.

@@

We have built a system---henceforth called OHM---which has the goal of being
a large scale aggregating data warehouse for time-series-based log-like
data sources.

@@

The user writes the map part of the MapReduce programs which imports
log-like data to the data store. The reduce step is taken care of by OHM,
which uses sum reducer that simply sums up values with the same keys.

@@

Import jobs use mappers defined by the user, either in Python or Java.
Mappers written by the user import each line from the input file and set
the arguments to the map function, key and value, as line number and line
contents, respectively. The user can then perform computations on these
entities and finally yield the results as a dictionary. The resulting
dictionary will be interpreted by OHM and stored accordingly.

\begin{verbatim}
   zohmg import mappers/somemapper.py path/on/hdfs
\end{verbatim}

@@

Data is stored in a BigTable like database. A natural consequence of this
that the data is stored sorted by rowkey allowing fast access to an
arbitrary (sequential) range.

@@

It is possible to query OHM about metadata it has on each of the imported
data sets.

Data is served to the user either over HTTP in JSON format or over Thrift.
\cite{thrift}

@@

\noindent The above serves on default port 8086, with the --port argument
it is possible to change server port.

To extract data, visit the server URL (e.g. http://localhost:8086/) and
give a request and query.

@@

Dispatching.

@@

Dispatches to static files in client directory in the project directory.
Custom clients can be built with HTML and JavaScript, see the examples.

@@

Serves data from HBase, dumps it as JSONP.

@@

Fetches data from HBase and transforms it with the requested transform
program in the transformers directory in the project directory, output is
dumped as JSONP.

@@

Update tables on configuration---dataset files---changes.

>>>>

Each column-family is stored as a single map-file in HBase.

@@

Each column-family is further divided into one or more column-qualifiers.

@@

Hadoop is a free software implementation of Google's MapReduce-framework
described in \cite{mapreduce}. Hadoop is written in Java and is part of
the Apache Software Foundation.  It is a fairly mature piece of software
and is used in production at numerous companies.

@@

MapReduce works by splitting the input, which commonly is in the size
range of many gigabytes, into n parts and lets each node of the computing
cluster work on one or more parts. The mapper on each node emits key-value
pairs which are fed into the reduce phase. The final output is collated,
usually on a distributed file system.

@@

HDFS is a free software implementation of the GFS \cite{gfs}.

The goal of HDFS is to store large data sets while being able to handle
hardware failure and being deployed in heterogenous hardware and software
eco systems. Additionally the Hadoop and HDFS interaction is designed with
the notion that it is cheaper to move the computation, MapReduce programs,
to the data than the other way around.

Every Hadoop node can be formatted with HDFS, this reserves disk space on
the node to be used for HDFS. The HDFS container resides on the nodes
general purpose file system. It is also possible to use HDFS as a
stand-alone general purpose distributed file system, without Hadoop.

The default block size on HDFS 64 MB, significantly larger than
file systems used for hard drives. The motivation is that applications
that use HDFS are not general purpose applications which run on general
purpose file systems, but batch applications which read, write, or both,
large amounts of data.

@@

Any program that reads and outputs to the file pointers \texttt{stdin} and
\texttt{stdout}, respectively, can be used as MapReduce programs with
Hadoop Streaming. This makes it possible to use any shell script or
program which inputs and outputs this way as a MapReduce program.

@@

HBase is a free software implementation of Google's BigTable data store
described in \cite{bigtable}. It is a sub-project of Hadoop, written in
Java and a part of the Apache Software Foundation. HBase is still in its
infancy and few organizations use it for mission-critical applications.

@@

TODO: *this should go into implementation.*
The typical use case for Zohmg is to read one or more column-qualifiers,
which correspond to one or more points in n-space, over a range of time.
Therefor, the keys are made up of the time and unit of the measurement,
and the qualifier queried for represents the point in n-space. More on
this in the data model section.

@@

@@

Python Setuptools is a framework---built on top of the Python module
distutils---for basic package management of Python software. Amongst other
things it creates and installs Python Eggs\footnote{A Zip file which
bundles Python software files.}.

@@

As stated in chapter 3, Zohmg is built upon existing tools and software.
It uses Apache Hadoop and Apache HBase, free software implementations of
Google's MapReduce and BigTable respectively, both implemented in Java.

@@

The resulting aggregated data, from the mapper, is stored in Apache HBase.

Exporting data from Zohmg is done via a web interface which dumps JSON.
It is possible to transform the stored data before exporting it with user
defined transformers.

Zohmg is implemented in Python.

@@

Certain paths need to be known by Zohmg in order to function, these are set
in the config/environment.py file.

@@

The user defines a mapper which Zohmg wraps, interpreting the output from the
mapper program and storing it in HBase.

@@

We have built a system---henceforth called Zohmg---which has the goal of being
a large scale aggregating data warehouse for time-series-based log-like
data sources.

@@

The user writes the map part of the MapReduce programs which imports
log-like data to the data store. The reduce step is taken care of by Zohmg,
which uses sum reducer that simply sums up values with the same keys.

@@

Import jobs use mappers defined by the user, either in Python or Java.
Mappers written by the user import each line from the input file and set
the arguments to the map function, key and value, as line number and line
contents, respectively. The user can then perform computations on these
entities and finally yield the results as a dictionary. The resulting
dictionary will be interpreted by Zohmg and stored accordingly.

\begin{verbatim}
   zohmg import mappers/somemapper.py hdfs/path
\end{verbatim}

@@

Data is stored in Apache HBase. A natural consequence of this that the
data is stored sorted by rowkey allowing fast access to an arbitrary
(sequential) range.

@@

It is possible to query Zohmg about metadata it has on each of the imported
data sets.

Data is served to the user either over HTTP in JSON format or over Thrift.

@@

\noindent The above serves on default port 8086, with the \texttt{--port}
argument it is possible to change server port.

To extract data, visit the server URL (e.g. http://localhost:8086/) and
give a query request.

@@

Middleware which dispatches to applications based on the URI.

@@

Dispatches request to serve static files from the project's client directory.
Custom clients can be built with HTML and JavaScript, see the examples.

@@

Serves data from HBase as JSONP\footnote{JSON with padding: JSON extension which adds name of callback function as an argument to the function itself.}.

@@

Fetches data from HBase and transforms it with the requested transform
program from the project's transformers directory, output is
dumped as JSONP.

@@

Update tables on changes in configuration---dataset files.
